---
title: "Assignment 4.1 Hill, Zach"
author: "Zach Hill"
date: "05APR2019"
output:
  html_document:
    df_print: paged
---

```{r setup, echo = FALSE, results = 'hide', message = FALSE}
library(rmarkdown)
library(knitr)
library(ggplot2)
library(readr)
library(pastecs)
library(psych) 
library(grid)
library(gridExtra)
exams <- read_tsv('Exams.dat')
exams$uni <- factor(exams$uni, levels = c(0:2), labels = c("George", "Jeff", "Tushmann"))
```

# Desciptive Statistics


```{r describe, echo = FALSE}
options(digits = 2)
stat.desc(exams[1:5], basic = FALSE)
by(exams[1:5], exams$uni, stat.desc, basic = FALSE)
```


```{r subsets, echo = FALSE}
georgeData <- subset(exams, exams$uni == "George")
jeffData <- subset(exams, exams$uni == "Jeff")
tushmannData <- subset(exams, exams$uni == "Tushmann")
```

## Professor George

Professor George's class tends towards lower scores in the exam and on numeracy as shown below.

```{r george, echo=FALSE, fig.width = 8, fig.height = 3, fig.show = 'hold', fig.align = 'center'}
ggplot(georgeData, aes(exam)) +
  geom_histogram(aes(y = ..density.., fill = uni), binwidth = 5) +
  stat_function(fun = dnorm, args = list(mean = mean(exams$exam), sd = sd(exams$exam)), colour = "black", size =1) -> fig1

ggplot(georgeData, aes(numeracy)) +
  geom_histogram(aes(y = ..density.., fill = uni), binwidth = 3) +
  stat_function(fun = dnorm, args = list(mean = mean(exams$numeracy), sd = sd(exams$numeracy)), colour = "black", size =1) -> fig2

grid.arrange(fig1, fig2, ncol = 2)
```

We can see that Professor Georges students' median scores were below the mean, skewing the histograms right. 

## Professor Jeff

Professor Jeff's class tends towards higher scores in the exam and on computer assessments as shown below.

```{r jeff, echo=FALSE, fig.width = 8, fig.height = 3, fig.show = 'hold', fig.align = 'center'}
ggplot(jeffData, aes(exam)) +
  geom_histogram(aes(y = ..density.., fill = uni), binwidth = 5) +
  stat_function(fun = dnorm, args = list(mean = mean(exams$exam), sd = sd(exams$exam)), colour = "black", size =1) -> fig3

ggplot(jeffData, aes(computer)) +
  geom_histogram(aes(y = ..density.., fill = uni), binwidth = 3) +
  stat_function(fun = dnorm, args = list(mean = mean(exams$computer), sd = sd(exams$computer)), colour = "black", size =1) -> fig4

grid.arrange(fig3, fig4, ncol = 2)
```

We can see here Professor Jeff's students' median scores were above the overall mean skewing the histograms left.

## Professor Tushmann

Professor Tushmann's class tends towards higher scores in the exam and on computer assessments as shown below.

```{r tushmann, echo=FALSE, fig.width = 8, fig.height = 3, fig.show = 'hold', fig.align = 'center'}
ggplot(tushmannData, aes(exam)) +
  geom_histogram(aes(y = ..density.., fill = uni), binwidth = 5) +
  stat_function(fun = dnorm, args = list(mean = mean(exams$exam), sd = sd(exams$exam)), colour = "black", size =1) -> fig3

ggplot(tushmannData, aes(computer)) +
  geom_histogram(aes(y = ..density.., fill = uni), binwidth = 3) +
  stat_function(fun = dnorm, args = list(mean = mean(exams$computer), sd = sd(exams$computer)), colour = "black", size =1) -> fig4

grid.arrange(fig3, fig4, ncol = 2)
```

We can see here Professor Tushmann's students' median scores were below the overall mean skewing the histograms right.

## Shapiro-Wilk test

As we can see, most of the samples are not normally distributed as indicated by the sub-.05 p-value with the exception of the computer assessment

```{r, echo = FALSE}
shapiro.test(exams$exam)
shapiro.test(exams$computer)
shapiro.test(exams$lectures)
shapiro.test(exams$numeracy)
shapiro.test(exams$stats)
```

## Shapiro-Wilk test by professor

Most sample distributions appear normal with the exception of the stats and numeracy, and Professor Jeff's computer assessments
```{r, echo = FALSE, results = 'hide'}
by(exams$exam, exams$uni, shapiro.test)
```



Normalcy of computer assessments
```{r, echo = FALSE}
by(exams$computer, exams$uni, shapiro.test)
ggplot(exams, aes(sample = exams$computer, position = 'jitter'), width = 3, height = 3) + stat_qq()
```

Professor Jeff's students shows significant results with a .02 p-value

```{r, echo = FALSE, results = 'hide'}
by(exams$lectures, exams$uni, shapiro.test)
```



```{r, echo = FALSE}
by(exams$numeracy, exams$uni, shapiro.test)
ggplot(exams, aes(sample = exams$numeracy, position = 'jitter'), width = 3, height = 3) + stat_qq()
```

```{r, echo = FALSE}
by(exams$stats, exams$uni, shapiro.test)
ggplot(exams, aes(sample = exams$stats, position = 'jitter'), width = 3, height = 3) + stat_qq()
```